\documentclass[a4paper,12pt]{article}

\usepackage{verbatim}
\usepackage{booktabs}
\usepackage{nameref}
\usepackage{tabularx}
\usepackage{caption}
\usepackage{gb4e}

\noautomath

\captionsetup[table]{}

\renewcommand{\familydefault}{\rmdefault}
\renewcommand*\rmdefault{bch}

\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}

\newcounter{magicrownumbers}
\newcommand\rownumber{\refstepcounter{magicrownumbers}\arabic{magicrownumbers}}

\begin{document}

\title{Explanatory Style Annotation Guide -- DRAFT}
\author{HiLT Lab}
\date{\today\\v0.5.2}
\maketitle

\tableofcontents
\break

Research suggests that the way that people explain things that happen in their lives can be revealing.\footnote{\label{seligman89}Schulman, P., Castellon, C., \& Seligman, M. (1989). Assessing explanatory style: The content analysis of verbatim explanations and the attributional style questionnaire. Behavior Research and Therapy, 27(5), 505-512} % TODO: try to find a more appropriate citation; preferably one that doesn't talk as much about the questionnaire (or perhaps just mention the questionnaire)
A consistently pessimistic \emph{explanatory style}, for example, can help indicate depression.
The ability to detect these explanations, and consequently describe a person's explanatory style, is therefore a valuable skill for therapy-minded Companionbots since it lets them know when and about what to start a therapeutic dialogue.

\section{Event Attribution Units (EAUs)}
In order to describe a person's explanatory style, we need a sample of how they explain things that happen to them (particularly things about which the person has a non-neutral opinion). % TODO: briefly describe CAVE analysis and difference from questionnaire
This sample is made up of a set of \emph{event-attribution units}\footnote{\label{CAVE}appendix of CAVE paper, with guidelines} (EAUs) and some descriptive features about them. % TODO: citation needed
EAUs are the conjunction of an \emph{event} and an \emph{attribution} that explains the existence of the event.


\subsection{Events}

Seligman provides a fairly comprehensive definition of what is meant by \emph{event} in his annotation guidelines: % TODO: cite, revise "his"

\begin{quote}
    An event is defined as any stimulus that occurs in an individual's environment or within that individual (e.g. thoughts or feelings) that has a good or bad effect from the individual's point of view.
    Events can be mental (e.g. I was afraid), social (e.g. I got a pay raise) or physical (e.g. I got in a car accident).
    Events should be unambiguously good or bad from the individual's point of view and may occur in the past, present or hypothetical future.
    Events that have good and bad elements, neutral events or events that do not affect the \emph{S} [the individual] should not be extracted.
\end{quote}

While we do attempt to follow this definition as closely as possible, we find that it is easier to reproduce similarly formed \emph{events} from our own transcripts by requiring that several conditions be true about our annotations in addition to those given in Seligman's guidelines. Those conditions are located in section \ref{conditions} on page \pageref{conditions}.

\pagebreak
\subsection{Attributions}
\subsection{Examples}

(...) He then goes on to give examples of good \emph{event-attribution units}: % TODO: reproduce examples
\setcounter{magicrownumbers}{0}
\begin{table}[h!]\centering
    \ra{1.3}
    \caption{Good EAUs.}
    \label{table1}
    \noindent\begin{tabularx}{\textwidth}{llX}
        \toprule
        & Event & Attribution \\
        \midrule
        \rownumber & I didn't do well on my exam & because I didn't sleep well last night. \\
        \rownumber\label{ex_grad_prog} & I haven't been sleeping well & because I'm worried about getting into a good graduate program. \\
        \bottomrule
    \end{tabularx}
\end{table}

%I want to refer to example \ref{table1}.\ref{ex_grad_prog} %example_graduate_program

\section{Analyzing EAU dimensions}
    \subsection{Personal v. External}
    \subsection{Permanent v. Temporary}
    \subsection{Pervasive v. Specific}
    \subsection{Examples}

\pagebreak
\section{Annotating Event-Attribution Units and their Features}

    \subsection{Annotation Phases}
    \subsubsection{Phase 1: Event spans}
    Two annotators, individually:
        \begin{enumerate}
            \item Annotate the most appropriate text span of each event they identify as part of an EAU as type \texttt{Event}.
        \end{enumerate}
    Consensus procedure:
        \begin{enumerate}
            \item Annotators discuss each given annotation and copy to the consensus set those for which they agree meet all given annotation conditions, including the existence of an associated attribution.
                \begin{enumerate}
                    \item Annotators may resolve to include any annotation (whether it have been made during individual annotation or not) in the consensus set that they may happen to notice during discussion which they deem appropriate to its definition.
                \end{enumerate}
        \end{enumerate}


    \subsubsection{Phase 1B: Event polarity}
    One annotator:
        \begin{enumerate}
            \item Annotates the polarity of each event given in the consensus set from Phase 1 as either \texttt{Positive} or \texttt{Negative} as a feature of the annotation.
        \end{enumerate}

    \subsubsection{Phase 2: Attribution spans}
    Two annotators, individually:
        \begin{enumerate}
            \item Annotate, for every \texttt{Event} annotation from Phase 1B, the text span of every associated attribution as type \texttt{Attribution}.
                \begin{enumerate}
                    \item Provide the ID number of the associated event as a feature, \texttt{Caused\_Event}.
                \end{enumerate}
        \end{enumerate}
    Consensus procedure:
        \begin{enumerate}
            \item Annotators discuss each given annotation and copy to the consensus set those for which they agree meet all given annotation conditions.
                \begin{enumerate}
                    \item Annotators may resolve to include any annotation (whether it have been made during individual annotation or not) in the consensus set that they may happen to notice during discussion which they deem appropriate to its definition.
                \end{enumerate}
        \end{enumerate}

    \subsubsection{Phase 3: Attribution dimensionality}
    For every attribution within the given consensus set:
        \begin{enumerate}
            \item Annotate values for each of the three dimensions as features, \texttt{Personal--External}, \texttt{Permanent--Temporary}, \texttt{Pervasive--Specific}, each with a digit betwen \texttt{1} and \texttt{7}.
        \end{enumerate}
    \noindent Some tips regarding the annotation of the three dimensions:
        \begin{enumerate}
        \item Internality
            \begin{description}
                \item[Who is to blame? Who takes the credit?] In answering the question, ``What degree of internality/externality is present in this Attribution?" you will necessarily be answering the question, ``According to the speaker, who/what is to blame (in the case of a negative event) or who/what should take credit (in the case of a positive event) for the associated Event?"
            \end{description}
        \item Stability
            \begin{description}
                \item[Temporally constrained Attributions, temporally abstracted Events] When annotating the stability of an Attribution, you are essentially annotating the likelihood that the Attribution will be considered a cause of the Event by the speaker if the Event were to happen again. In answering this question, it will be necessary to hypothesize about future occurences of the Event, but be careful not to extend this hypothetical way of thinking to the Attribution itself. That is to say that in answering this question, your mental model should consider the Attribution constrained to the particular instance mentioned in time, but consideration of the Event should be abstracted to all possible future occurences of the Event, regardless of how unlikely it really would be to happen again.
            \end{description}
        \item Globality
            \begin{description}
                \item[Avoid considering butterfly effects] When faced with the task of mentally modeling the average person as Seligman suggests in his guidelines in the case of a lack of knowledge of the subject's opinions in answering the question, ``How many domains of the speakers life is likely to be believed to be affected?" try only to count obviously immediately following effects and not all of the likely chain reactions of each of those effects (think: butterfly effects). That is to say, if given an Attribution to the effect of ``they're changing my (blood pressure) medication," the only obvious domain in which that plays a major role is the speaker's physical health, but a large range of future changes might possibly or even probably change as a result. However, the majority of these further effects would likely only be a result of whatever physical change the medication brings about, and not the change in medication itself. All this isn't to say, however, that an Attribution with an otherwise narrow scope of effect could never be annotated as more pervasive; if you believe the speaker believes it's pervasive, mark it as such.
            \end{description}
        \end{enumerate}

    \newpage
    \subsection{Annotation span conditions} \label{conditions}
        Some definitions:
        \begin{description}
            \item[propose, \emph{v.}] to make a statement about something which could be evaluated as either \emph{true} or \emph{false},
                e.g. \emph{a bell rang repeatedly}, but not just \emph{a bell} or \emph{to ring} or \emph{repeatedly}.
            \item[corefer, \emph{v.}] to share a single reference,
                e.g. \emph{Paul} and \emph{He} in ``\emph{Paul} had a great time" and ``\emph{He}'s always in a good mood."
            \item[backchannel, \emph{v.}] to affirm active listening, e.g. ``okay," ``uh-huh," ``I'm listening," etc.
            \item[turn \emph{n.}] in conversation, an uninterrupted span of speech by a single person (excepting any \emph{backchannel}).
        \end{description}
    \subsubsection{Events}
    For every \textbf{\emph{event}} annotation:
        \begin{enumerate}
            \item The \emph{event} must have been mentioned during the given conversation.
            \item The event must have been or otherwise be very probable to have been evaluated as either good or bad by subject in terms of its effects on the subject.
            \item The text span that expresses the \emph{event} must be in the form of a complete sentence or otherwise unambiguously \emph{corefer} with a complete sentence present in the transcript expressing such event when no other non-pronoun mention of the event appears within the same \emph{turn} as the otherwise most appropriate event mention.
            \item The text span that expresses the \emph{event} must primarily \emph{propose} that event. %...", and nothing else."? % TODO: give examples as reasons
            \item Experience of the \emph{event} by the participant must not rely solely on a generic statement.
                e.g. not ``Teachers often feel stressed'' even if the individual is in fact a teacher.
            \item The text span expressing the \emph{event} must be the minimal length necessary to satisfy all other conditions.
        \end{enumerate}


    \subsubsection{Attributions}
    For every \textbf{\emph{attribution}} annotation:
        \begin{enumerate}
            \item The \emph{attribution} must have been mentioned during conversation.
            \item The text span that expresses the \emph{attribution} must primarily \emph{suggest} that attribution. %...", and nothing else."?
            \item \emph{Events} must be the minimal length necessary to satisfy all other conditions.
            \item The causal relationship identified between the attribution and its associated event must have clearly been intended to have been communicated.
            \item The internality dimension of the attribution must be non-neutral (i.e. either internal or external).
        \end{enumerate}


    \subsubsection{All annotations}
    For \textbf{every annotation} made:
        \begin{enumerate}
            \item Annotations may be non-contiguous.
                \begin{enumerate}
                    \item Use \emph{annotation}\texttt{\_continuation}, replacing \emph{annotation} with the type name of the continued annotation, e.g. \texttt{Event\_continuation}.
                    \item Annotation continuations must not be interrupted by a complete annotation of the same type.
                \end{enumerate}
            \item Annotation spans may fail to provide all relevant information and remain valid only if nearby contextual information relieves the insufficiency.
            \item Avoid annotating sentence-ending punctuation at annotation boundaries.
        \end{enumerate}


    \subsubsection{EAUs}
    \textbf{Every \emph{EAU}}:
        \begin{enumerate}
            \item must consist of the nearest pair of annotations for which both spans satisfy their respective conditions, except:
                \begin{enumerate}
                    \item when an attribution is repeated in such a way that does not seem to be merely sentence repair (e.g. stuttering, clarification, etc.)
                    % TODO: start a "common mistakes" section and put this in it -> \item When the same \emph{event} is mentioned multiple times, for which an \emph{attribution} is given only once, only the nearer text span will be included, \emph{provided} the nearer span does not significantly rely on contextual information to meet it's other conditions.
                \end{enumerate}
            \item must not overlap in its event and attribution spans.
            \item given a contiguous series of similar events which an attribution or series of attributions with similar internality explains, annotate a single event attribution grouping like events as a single event and like attributions as a single attribution.
        \end{enumerate}

\end{document}
